Letter to the Editor

### 2022-04-21 editorial-communication/cover-letter-submission.Rmd
Deploying to gh-pages from @ SORTEE-Github-Hackathon/manuscript@cf98dfc0a509af73955fdbcf6b1a324c394838cf 🚀

We would like to submit our manuscript entitled “Not just for programmers: A friendly guide on the benefits of GitHub for accelerating collaborative research in Ecology and Evolution” for consideration to be published in *Nature Ecology & Evolution*. 
Our study provides researchers in ecology and evolution (E&E) with a critical review of thirteen practical ways they can use Github -- an online, cloud-based service for cooperative and documented code development -- to make their research more collaborative, transparent, and reproducible.

Scientific advancement fundamentally requires researchers to adopt reproducible, collaborative and transparent practices.

Most scientists, including ecologists and evolutionary biologists, have become critically dependent on computational tools in their research.
This code-dependent workflow imposes steep requirements towards an efficient, well-documented process in the publication and collaboration of maintainable scientific code.
Albeit, many researchers lack exposure to adequate software development practices and thus are required to dedicate valuable time and effort to self-teach the use of research-facilitating tools, and thus may find practical barriers when applying adequate standards to maintain their scientific code.
The adoption of GitHub in E&E workflows is not widespread in part because there have been only a few disparate attempts at creating domain-specific information and guidelines.
In our study, we review and discuss one of the most used web-based platforms for computational version control and collaboration, GitHub, and provide researchers in ecology and evolutionary biology with practical workflows aimed at facilitating their scientific code and management process.

Our manuscript is the most accessible and practical guide to using GitHub in E&E research to date. 
This is because we represent a wide diversity of E&E domains, yet have all found GitHub essential to making our research more collaborative and transparent. 
Our manuscript is important for the *Nature Ecology & Evolution* readership because EEB researchers will continue to migrate their workflows to GitHub and need resources to guide their journey toward more open and collaborative EEB science.

The enclosed manuscript has not been previously published and is not under consideration for publication in another journal or book. 
All authors have contributed and reviewed the current version of the manuscript and agreed to its submission to *Nature Ecology & Evolution*.

>>>>

We would like to submit our manuscript entitled “Not just for programmers: A friendly guide on the benefits of GitHub for accelerating collaborative research in Ecology and Evolution” for consideration to be published in *Nature Ecology & Evolution*. 
Our study provides ecology and evolutionary biology (EEB) researchers with a critical review of thirteen practical ways they can use Github -- an online, cloud-based service for cooperative and documented code development -- to make their research more collaborative, transparent, and reproducible.

Most scientists, including researchers in EEB, have become critically dependent on computational tools in their research.
This code-dependent workflow imposes steep requirements towards an efficient, well-documented process in the publication and collaboration of maintainable scientific code.
Albeit, many researchers lack exposure to adequate software development practices and thus are required to dedicate valuable time and effort to learning how to use research-facilitating tools. 
Formal academic courses on these types of tools in EEB are often unavailable and researchers are left to learn through informal blogs, videos, and other online forums. 
This may present a plethora of practical barriers to applying adequate standards to maintaining scientific code.

The adoption of GitHub in EEB workflows is not widespread in part because there have been only a few disparate attempts at creating domain-specific information and guidelines.
In our study, we review and discuss one of the most used web-based platforms for computational version control and collaboration, GitHub, and provide researchers in EEB with practical workflows aimed at facilitating their scientific code and management process.

Our manuscript is the most accessible and practical guide to using GitHub in EEB research to date. 
We build upon some existing resources, which are mainly targeted at other disciplines, and our own experiences (from across many EEB subdisciplines) in writing and sharing code openly, to highlight steps EEB researchers can take to improve their research practices in a reproducible, collaborative, and transparent way.
Our manuscript is important for the *Nature Ecology & Evolution* readership because EEB researchers will continue to migrate their workflows to GitHub and need resources to guide their journey toward more open and collaborative science.

The enclosed manuscript is available as a preprint on the NAME OF SERVER AND CITATION, but is not published or in review elsewhere.
All authors have contributed and reviewed the current version of the manuscript and agreed to its submission to *Nature Ecology & Evolution*.
